Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-08-17 manuscript/revision2/response_to_reviews/response_to_reviews.tex

We were pleased by the useful feedback given by the reviewers of our
original submission and believe that the greedy mash algorithm we
implemented addresses the major statistical concern of reviewer 2, as
well as substantially improving our use and interpretation of covariance
matrices in \emph{mash}. We have addressed all the editorial issues
suggested by the two reviewers, and included in this letter are
point-by-point responses to these issues.

We suggest that the original editor, Editorial Board members, and
anonymous reviewers for this manuscript would be good choices to assess
this revision. In the event that is not practicable, Douglas Schemske,
Stephen P. Long, and Qifa Zhang would be good Editorial Board members
for this manuscript. All would be able to provide strong insights into
mapping the genetic basis of adaptive and ecologically relevant traits
in a field experimental context. Johanna Schmitt, Trudy Mackay, and
Detlef Weigel would be excellent editors for this manuscript.
Dr.~Schmitt has done extensive work in \emph{Arabidopsis thaliana} and
other plant species unravelling the genetic mechanisms involved in
responses to the natural environment. Dr.~Mackay has extensive
experience mapping the genetics of complex traits and
genotype-by-environment interactions. Dr.~Weigel has done extensive work
understanding genetic diversity in \emph{A. thaliana} including
extensive work on floral induction cues.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

Reviewer \#1:\\
\strut \\
Suitable Quality?: No\\
Sufficient General Interest?: No\\
Conclusions Justified?: Yes\\
Clearly Written?: No\\
Procedures Described?: Yes\\
\strut \\
Comments on Significance Statement:\\
\strut \\
Conclusions are very specific for the crop/population and traits.
General conclusions are lacking.\\
\strut \\
Comments:\\
\strut \\
The paper by MacQueen et al.~presents results of a study on two
switchgrass populations grown in 8 garden experiments across a rane of
latitudes. Traits are green-up date and flowering time. The data show
strong GxE, effects of QTL changing in magnitude and sign. Results are
partly validated in an independent segregating population.\\
\strut \\
The topic is timely and highly relevant. A better understanding of GxE
will be valuable for plant genetics, evolution and breeding. The data
are complex and the authors have succeeded in making them accessible to
the reader.\\
\strut \\
\strut \\
Relevant literature\\
The introduction and discussion are focused on findings in Arabidopsis
and switchgrass. An excellent study from Arabidopsis with high relevance
for this paper is\\
Fournier-Level et al.~2016
\url{https://doi.org/10.1073/pnas.1517456113}\\
\strut \\
Major findings from other crop species have been ignored. To mention
just two references:\\
\strut \\
Bustos-Korts et al.~2019 The Plant Journal
\url{https://doi.org/10.1111/tpj.14414}\\
Millet et al.~2016 Plant Physiology
\url{https://doi.org/10.1104/pp.16.00621} and\\
There are many papers on crop modelling (eg work by Mark Cooper and
others in maize) and the authors might want to consider linking their
results to crop modelling to demonstrate the relevance of the results of
this study

\end{tcolorbox}
\end{quote}

We now add these citations (lines 51-54; lines 62, 125-131). We regret
we can't comprehensively reference the impressive crop modeling
literature. Our goal was to reference work on natural populations that
focused the genetics on phenology and/or local adaptation, and to touch
on how these ideas have interacted with crop improvement. This was
because our work uses a natural population of switchgrass with potential
relevance for breeding efforts in this species.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

Results\\
\strut \\
There is a large overlap of data and findings with reference 32 Lovell
et al.~2021 Nature. It has not become clear which additional insights
can be gained from this study. One reason might be that objectives of
the study have remained rather vague as well as the generic conclusions
that can be drawn from the results.

\end{tcolorbox}
\end{quote}

While we do rely on the same genotypic data and common gardens used in
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021}, and make use of the population structure findings of
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021} to divide our genotypes into subpopulations for GWAS, the previous
work considered fitness proxies and this work considers phenological
traits. This work has two additional, key advances over the previous
work: it can assign loci genome-wide to multiple kinds of GxE and
GxWeather, and has an unbiased statistical method to identify loci with
and without rank-changing GxE. We now summarize the findings from
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021} that we use as a springboard for this work on lines 142-150, and
outline the additional advances of this study on lines 162-169, 175-177,
and 884-895.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

In the introduction the aim is given as „we test if these populations
differ in their phenological adaptation and hence their phenological
GxE''. Given prior knowledge such as the origin of the populations and
earlier results (Lovell et al.) what is expected?

\end{tcolorbox}
\end{quote}

We now edit the Introduction to more clearly articulate the aims of this
manuscript. Our key expectation is that different genetic subpopulations
and genomic regions have likely evolved distinct patterns of GxE (lines
59-62; lines 228-232). Thus, we aim to identify the kinds of GxE present
for each subpopulation and each trait. (lines 135-137).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

The results for these two populations did not show a general pattern
leading to conclusions on the environmental cues. Associations trait/cue
changed across populations, traits, gardens, analysis etc. , it was
difficult to see the big picture.

\end{tcolorbox}
\end{quote}

We agree that we do not identify one general pattern or one
environmental cue affecting any set of populations or phenological
trait. Instead, our key expectation is that different populations and
different genetic loci will have evolved distinct patterns of GxE (lines
59-62; lines 228-232; SI Appendix Section S1). Thus, we aim to identify
the kinds of GxE present genome-wide and elaborate when these kinds of
GxE differ (Figure 2), and demonstrate that individual loci have
different types of GxE across environmental regions and subpopulations
(Figure 3).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

Results from the SNPs ``mash model of Midwest green-up fell on a
covariance matrix of average temperature in the 10 days prior to Midwest
green-up''. These results should be linked to the phenotypic GxE
analysis.

\end{tcolorbox}
\end{quote}

We edit the results to directly compare the GxWeather matrices and the
phenotypic correlations in Figure 1B. (lines 251-269; lines 284-287;
lines 290-293; lines 297-301).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

Discussion\\
``we must understand the current patterns of trait covariation across
environments, \ldots..''. What is the understanding from your data? Are
there any general patterns or do we need to assess them individually for
different genetic material, environments, traits? It looks like the
latter so what are the consequences?

\end{tcolorbox}
\end{quote}

As we state in the SI Appendix (Section S1, paragraph two), we were
interested in the types of genetic correlation that could be expressed
by different alleles across the genome. We reasoned that different loci
could have different patterns of genetic correlation. For example, the
effects of an allele of locus \(A\) could covary with a weather-based
cue if \(A\) had a common allele, \(a\), that was responsive to this
cue. In contrast, the effects of a different locus \(B\) might have an
effect at only one garden, for example if a pathogen was only present at
one garden, and \(B\) has a common allele \(b\) that is resistant to
that pathogen.

As a consequence, we develop an approach to specify multiple
environmental cues and compete them to explain patterns of genetic
effects (lines 238-244): we use a greedy algorithm to use with mash to
select covariance patterns, and use mash to flexibly identify these
patterns across populations and environments (lines 244-250). With this
approach, we are able to demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions (lines 427-430). We
can also assign genetic effects to both GxWeather patterns derived from
weather variables and to other, site-based patterns (lines 430-432).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

``Gulf and Midwest subpopulations have two distinct photoperiod-related
flowering responses\ldots.''. This has been shown for other species, eg
Unterseer et al.~2016 Genome Biology

\end{tcolorbox}
\end{quote}

We now reference this paper in the introduction (lines 51-56).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

The reader should be able to understand what the hypothesis in the
hypothesis-driven models is without looking at the supplement.

\end{tcolorbox}
\end{quote}

Agreed. We add Table 1 to better explain the covariance matrices we
select using the greedy mash algorithm. We include the matrices selected
by this algorithm that have \textgreater0.1\% weight (lines 273-274) in
the mash model as visualizations in Figure 2.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

@@

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

Statistical issues:\\
\strut \\
- Narrow-sense heritability: A key result is that the h\^{}2 is low when
measured across all trials, but high within trials. However the model
used to estimate the global h\^{}2 is missing a term for replication of
lines across accessions. u only accounts for replication of additive
effects of lines across gardens, but non-additive genetics can be
persistent too. Also, this model doesn't allow var(e) to vary across
locations but it's likely that it does. If the goal is to show
rank-crossing GxE, actually fit a GxE model like was done for the
environmental indices to directly show it.\\
- Variance components analysis: The specification of this model is
incorrect because Var(u) and Var(ul) are assigned the same distribution.
There are many ways of specifying GxE effects in a model like this, and
it's not clear how it is done here. If G is a nxn GRM for n accessions,
it can't be the covariance matrix for both Var(u) (dimension nxn) and
Var(ul) (dimension nl x nl). Generally, Var(ul) would be G \otimes Psi
where Psi is a lxl covariance matrix among gardens. Sometimes this is
constrained to be diagonal (ie no covariance among gardens), and
sometimes to be proportional to I (constant variance and no covariance
among gardens), but these more restrictive models should be justified.

\end{tcolorbox}
\end{quote}

We agree that correctly specifying the type of GxE in our models of
additive variation across gardens is important for all models in this
manuscript. In many cases the complexity of the genetic
variance-covariance structures we wish to fit are not be possible to fit
using standard approaches. Eight gardens give eight variance components
and 8 x 8 covariances - most datasets are not large enough to give
solutions for these models. The standard approach is to either do data
reduction (e.g.~factor analysis) or impose unrealistic constraints
(e.g., equal variances, or only positive covariances) in order to solve
these models. Our approach and our focus on mash was to allow us to fit
less structured variance-covariance structures, as well as more of them.

@@

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]

- Antagonistic Pleiotropy: Maybe I'm missing the idea of this analysis,
but it doesn't seem to me that this strategy for partitioning loci
between antagonistic pleiotropy and differential sensitivity would have
equal power to detect each class. It seems that antagonistic pleiotropy
should be a subset of differential sensitivity because the latter is
defined as simply a change in magnitude. But even if it's restricted to
``different magnitude but same sign'', the practical definition here
seems to be that antagonistic pleiotropy is detected when the signs are
different and both lfsrs \textless{} threshold (0.05?), while
differential sensitivity additionally requires a threshold on the
difference in effect sizes (0.4) with no justification for why this size
was chosen. This additional criterion will necessarily make the rate of
detection different. Even without this, with differential sensitivity
one effect size must be much larger than the other in absolute value
(which isn't the case with antagonistic pleiotropy), so differential
sensitivity loci will be more likely to pass the lfsr loci than
antagonistic pleiotropy loci. Note that previous literature here tried
to separate antagonistic pleiotropy from conditional neutrality (ie
where in one location the effect was zero), while here the contrast is
with differential sensitivity where effects are non-zero in all
locations. I'm not aware of previous discussions in the literature of
trying to differentiate these two specific classes of GxE variants and
not really sure theoretically what the importance would be.

\end{tcolorbox}
\end{quote}

We believe that the use of the lfsr to detect effects with rank-changing
GxE is a key advance in our manuscript, and so we have made changes to
the Introduction (lines 175-177), Results (lines 338-343), and Materials
\& Methods (lines 959-982; lines 996-998; lines 1055-1057) to explain
and motivate this statistical change.

@@

Our expectation is that in nature, effects will always differ.
Statistically, this difference will not always be detectable. Thus,
rather than ask ``Are these two effects different?'' - as we reasonably
expect two effects to be, even if this difference cannot be measured -
the local false sign rate answers a more meaningful question: Can we be
confident in the sign of this effect?

We think that if evolutionary geneticists are interested in antagonistic
pleiotropy and rank-changing GxE, then they should use the lfsr to
measure their confidence in the sign of the effect, rather than using
statistical tests where the null hypothesis is that the effect is
different than zero. However, if we use the lfsr, comparisons between
antagonistic pleiotropy and conditional neutrality are no longer
sensible, as we are no longer doing a statistical test that can robustly
detect conditional neutrality (just as the FDR does not robustly detect
antagonistic pleiotropy). This is justified because detection of
antagonistic pleiotropy is more important than detection of conditional
neutrality for models of local adaptation.

It's true that the threshold for differential sensitivity is arbitrary.
We have reworded the manuscript to stress that there is equal power to
detect effects of different sign as there are effects of the same sign
(lines 996-998; lines 1055-1057). Another class of effects we have in
our analysis that differ from classic antagonistic pleiotropy \&
conditional neutrality comparisons are effects which are not
distinguishable by sign or magnitude. It is not accurate that
differentially sensitive loci will be more likely to pass the lfsr,
because the loci are first selected by significant lfsr, then divided
further into effects that can \& cannot be distinguished by magnitude.
Thus effects that are significant, but have small magnitude in both
conditions will be labeled as not being distinguishable - essentially,
these effects do not have detectable GxE. Effects without GxE may be
even more common and should also be quantified. We propose that, if our
intent is to detect sign-changing GxE, then we should use an unbiased
statistical test to look at loci with and without sign-changing GxE, the
lfsr.
